### 2022-04-27 exploratory/20220426_main_plot_t_test.Rmd
made a table for t-tests

---
title: "20220426_main_plot_t_test"
author: "Yurun (Ellen) Ying"
date: "4/26/2022"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
# load libraries
library(tidyverse); library(ggplot2); library(wesanderson); library(magrittr); library(here); library(knitr); library(broom); library(effectsize)

# read file
casual_tib <- here("data/processed/casual_sex_sim.csv") %>% 
  read_csv() %>% 
  mutate(
    diff_likelihood = as_factor(diff_likelihood),
    diff_standard = as_factor(diff_standard),
    homosexual = as_factor(homosexual)
  )
```

## Main plots
- Violin plots of the diff x diff condition
	- grouped by gender
	- paneled by experiments
	- separate for two dependent variables: avg num of experiences (all), avg num of partners (all)

```{r violin_exp}
casual_tib %>% 
  filter(diff_likelihood == 1, diff_standard == 1) %>% 
  select(homosexual, m_exp_all, f_exp_all) %>% 
  pivot_longer(c(m_exp_all, f_exp_all), values_to = "exp_all",
               names_to = "gender", names_pattern = "(.)_exp_all") %>% 
  ggplot(aes(x = gender, y = exp_all)) +
  geom_violin(aes(fill = gender)) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  scale_fill_discrete(
    type = wes_palette("Darjeeling2", 2, type = "discrete"),
    name = "", labels = c("Women", "Men")
  ) +
  labs(x = "", y = "Average number of\nshort-term mating experiences",
       title = " Main plot 1: violin plot of average number of\nshort-term mating experiences") +
  facet_wrap(
    ~ homosexual,
    labeller = labeller(
      homosexual = c(
        `0` = "Experiment 1\nHeterosexual individuals",
        `1` = "Experiment 2\nGay men and lesbian women")
      )) +
  theme_minimal() +
  theme(
    axis.line = element_line(size = 0.3),
    axis.ticks.y = element_line(size = 0.3),
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    panel.spacing = unit(0, "npc"),
    strip.text = element_text(size = 10)
    )

```

```{r violin_partner}
casual_tib %>% 
  filter(diff_likelihood == 1, diff_standard == 1) %>% 
  select(homosexual, m_partner_all, f_partner_all) %>% 
  pivot_longer(c(m_partner_all, f_partner_all), values_to = "partner_all",
               names_to = "gender", names_pattern = "(.)_partner_all") %>% 
  ggplot(aes(x = gender, y = partner_all)) +
  geom_violin(aes(fill = gender)) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  coord_cartesian(ylim = c(0, 3)) +
  scale_fill_discrete(
    type = wes_palette("Darjeeling2", 2, type = "discrete"),
    name = "", labels = c("Women", "Men")
  ) +
  labs(x = "", y = "Average number of\nshort-term patners",
       title = " Main plot 2: violin plot of average\nnumber of short-term partners") +
  facet_wrap(
    ~ homosexual,
    labeller = labeller(
      homosexual = c(
        `0` = "Experiment 1\nHeterosexual individuals",
        `1` = "Experiment 2\nGay men and lesbian women")
      )) +
  theme_minimal() +
  theme(
    axis.line = element_line(size = 0.3),
    axis.ticks.y = element_line(size = 0.3),
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    panel.spacing = unit(0, "npc"),
    strip.text = element_text(size = 10)
    )
```

## T-tests
- Independent samples *t*-test between men and women, separate for conditions and experiments
- Output into tables with descriptives and test results, separate by variables

```{r t_all_var}
# create a long tibble
casual_long <- 
  casual_tib %>%
  unite(homosexual, diff_likelihood, diff_standard,
        col = "homosexual_dlikelihood_dstandard", sep = "_") %>% 
  pivot_longer(-c(seed, homosexual_dlikelihood_dstandard), 
               names_to = c("gender", ".value"), 
               names_pattern = "(.)_(.*)"
  )

# descriptives
casual_des <- 
  casual_long %>% 
  group_by(homosexual_dlikelihood_dstandard, gender) %>% 
  summarise(across(inpool:partner_ip, 
                   list(mean = mean, sd = sd),
                   .names = "{.col}.{.fn}")) %>% 
  pivot_wider(names_from = gender, 
              values_from = -c(homosexual_dlikelihood_dstandard, gender),
              names_glue = "{gender}_{.value}")

# t test by groups for exp_all variable
casual_t_test <- 
  casual_long %>% 
  nest_by(homosexual_dlikelihood_dstandard) %>% 
  mutate(
    inpool = t.test(inpool ~ gender, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    outpool = t.test(outpool ~ gender, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    exp_all = t.test(exp_all ~ gender, data = data) %>% tidy() %>% 
      # somehow the map function doesn't work here and I wonder why...
      select(statistic, p.value, parameter),
    exp_ip = t.test(exp_ip ~ gender, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    partner_all = t.test(partner_all ~ gender, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    partner_ip = t.test(partner_ip ~ gender, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter)
    ) %>% 
  select(-data) %>% 
  unnest(cols = -homosexual_dlikelihood_dstandard, names_sep = ".")

# effect size
casual_d <- 
  casual_long %>% 
  nest_by(homosexual_dlikelihood_dstandard) %>% 
  mutate(
    inpool.d = cohens_d(inpool ~ gender, data = data) %>% pull(Cohens_d),
    outpool.d = cohens_d(outpool ~ gender, data = data) %>% pull(Cohens_d),
    exp_all.d = cohens_d(exp_all ~ gender, data = data) %>% pull(Cohens_d),
    exp_ip.d = cohens_d(exp_ip ~ gender, data = data) %>% pull(Cohens_d),
    partner_all.d = cohens_d(partner_all ~ gender, data = data) %>% pull(Cohens_d),
    partner_ip.d = cohens_d(partner_ip ~ gender, data = data) %>% pull(Cohens_d)
    )

# combine the descriptives, test results, and effect size
full_join(casual_des, casual_t_test, casual_d, by = "homosexual_dlikelihood_dstandard") %>% 
  kable()
```