### 2022-05-02 exploratory/20220430_hetero_vs_gay.Rmd
added analyses to compare heterosexual and gay men, close #10

---
title: "20220430_hetero_vs_gay"
author: "Yurun (Ellen) Ying"
date: "4/30/2022"
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
# load libraries
library(tidyverse); library(ggplot2); library(wesanderson); library(magrittr); library(here); library(knitr); library(broom); library(effectsize)

# read file
casual_tib <- here("data/processed/casual_sex_sim.csv") %>% 
  read_csv() %>% 
  mutate(
    diff_likelihood = as_factor(diff_likelihood),
    diff_standard = as_factor(diff_standard),
    homosexual = as_factor(homosexual)
  )
```

## Plots
- Violin plots of the diff x diff condition
	- grouped by sexual orientation
	- paneled by dependent variables: avg num of experiences (all), avg num of partners (all)

```{r violin_plot, fig.height=8, fig.width=5}
y_lab <- labeller(
      variable = c(
        `exp_all` = "Average number of\nshort-term mating experiences",
        `partner_all` = "Average number of\nshort-term patners"))

casual_tib %>% 
  filter(diff_likelihood == 1, diff_standard == 1) %>% 
  select(homosexual, m_exp_all, m_partner_all) %>% 
  pivot_longer(c(m_exp_all, m_partner_all), values_to = "value",
               names_to = "variable", names_pattern = "m_(.*)") %>% 
  ggplot(aes(x = homosexual, y = value)) +
  geom_violin(aes(fill = homosexual)) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  scale_fill_discrete(
    type = wes_palette("Darjeeling2", 2, type = "discrete"),
    name = "", labels = c("Heterosexual men", "Gay men")
  ) +
  labs(x = "", y = NULL,
       title = " Violin plot of average number of short-term mating\nexperiences and  short-term mates, comparing\nheterosexualand gay men") +
  facet_wrap(
    ~ variable, ncol = 1, scales = "free",
    strip.position = "left",
    labeller = y_lab
    ) +
  theme_minimal() +
  theme(
    axis.line = element_line(size = 0.3),
    axis.ticks.y = element_line(size = 0.3),
    axis.text.x = element_blank(),
    panel.grid = element_blank(),
    panel.spacing = unit(0.5, "line"),
    strip.text = element_text(size = 11),
    strip.placement = "outside",
    legend.position = "bottom",
    legend.box.spacing = unit(0.1, "line")
    )

```

## T-tests
- Independent samples *t*-test between heterosexual and gay men on all outcome variables, separate for conditions
- Output into tables with descriptives and test results, separate by variables

```{r t_all_var}
# create a long tibble
casual_long <- 
  casual_tib %>%
  unite(diff_likelihood, diff_standard,
        col = "dlikelihood_dstandard", sep = "_") %>% 
  select(-c(seed, starts_with("f_")))

# descriptives
casual_des <- 
  casual_long %>% 
  group_by(dlikelihood_dstandard, homosexual) %>% 
  summarise(across(m_inpool:m_partner_ip, 
                   list(mean = mean, sd = sd),
                   .names = "{.fn}.{.col}"),
            .groups = "keep") %>% 
  pivot_wider(names_from = homosexual, 
              values_from = -c(dlikelihood_dstandard, homosexual),
              names_glue = "{.value}_{homosexual}")

# t test by groups for exp_all variable
casual_t_test <- 
  casual_long %>% 
  nest_by(dlikelihood_dstandard) %>% 
  mutate(
    m_inpool = t.test(m_inpool ~ homosexual, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    m_outpool = t.test(m_outpool ~ homosexual, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    m_exp_all = t.test(m_exp_all ~ homosexual, data = data) %>% tidy() %>% 
      # somehow the map function doesn't work here and I wonder why...
      select(statistic, p.value, parameter),
    m_exp_ip = t.test(m_exp_ip ~ homosexual, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    m_partner_all = t.test(m_partner_all ~ homosexual, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter),
    m_partner_ip = t.test(m_partner_ip ~ homosexual, data = data) %>% tidy() %>% 
      select(statistic, p.value, parameter)
    ) %>% 
  select(-data) %>% 
  unnest(cols = -dlikelihood_dstandard, names_sep = ".")

# effect size
casual_d <- 
  casual_long %>% 
  nest_by(dlikelihood_dstandard) %>% 
  mutate(
    m_inpool.d = cohens_d(m_inpool ~ homosexual, data = data) %>% pull(Cohens_d),
    m_outpool.d = cohens_d(m_outpool ~ homosexual, data = data) %>% pull(Cohens_d),
    m_exp_all.d = cohens_d(m_exp_all ~ homosexual, data = data) %>% pull(Cohens_d),
    m_exp_ip.d = cohens_d(m_exp_ip ~ homosexual, data = data) %>% pull(Cohens_d),
    m_partner_all.d = cohens_d(m_partner_all ~ homosexual, data = data) %>% pull(Cohens_d),
    m_partner_ip.d = cohens_d(m_partner_ip ~ homosexual, data = data) %>% pull(Cohens_d)
    ) %>% 
  select(-data)

# combine the descriptives, test results, and effect size
des_t_d <- 
  full_join(casual_des, casual_t_test, by = "dlikelihood_dstandard") %>% 
  full_join(., casual_d, by = "dlikelihood_dstandard")

# display the results 
des_t_d %>% kable()
```