### 2022-05-13 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\newcommand{\bcenter}{\begin{center}}
\newcommand{\ecenter}{\end{center}}
\newcommand{\btitlepage}{\begin{titlepage}}
\newcommand{\etitlepage}{\end{titlepage}}
\usepackage{setspace}\doublespacing
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{booktabs}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\begin{titlepage}

\begin{center}

\vspace*{30mm}

\hypertarget{an-agent-based-model-of-gender-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}{%
\section*{An Agent-based Model of Gender Differences in Short-term
Mating Behaviors as a Result of Mating
Preferences}\label{an-agent-based-model-of-gender-differences-in-short-term-mating-behaviors-as-a-result-of-mating-preferences}}
\addcontentsline{toc}{section}{An Agent-based Model of Gender
Differences in Short-term Mating Behaviors as a Result of Mating
Preferences}

\vspace{30mm}

Yurun Ying\textsuperscript{1}, Jan Antfolk\textsuperscript{2}, Pekka
Santtila\textsuperscript{1,3,\(\dagger\)}

\textsuperscript{1} Faculty of Arts and Sciences, NYU Shanghai

\textsuperscript{2} xxxx

\textsuperscript{3} NYU-ECNU Institute for Social Development at NYU
Shanghai

\end{center}

\vspace{40mm}

\textsuperscript{\(\dagger\)} To whom correspondence should be
addressed: \url{pekka.santtila@nyu.edu}

\end{titlepage}

\newpage

\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

Gender differences in short-term mating behaviors have been a
well-documented phenomenon in human sexuality research. Existing studies
usually conflate gender differences in mating preferences with
differences in sexual behaviors, which is theoretically dubious. Using
an agent-based model, the present study investigated the circumstances
under which men and women's differential preferences for short-term
mating would result in gender differences in short-term mating
behaviors. The results from the model suggested that heterosexual men
and women had the same average number of short-term mating experiences
and short-term mates even when men had stronger preferences for
short-term mating. Gender differences in mating behaviors only emerged
when heterosexual men and women in the mating pool (i.e., those with a
non-zero number of short-term mating experiences) were considered.
Moreover, gay men had a higher average number of both outcomes than
lesbian women as well as than heterosexual men when men had stronger
preferences for short-term mating. These results suggest that
theoretically speaking, even when gender differences in mating
preferences existed, those in short-term mating behaviors only occurred
among particular populations, or when men's preferences for short-term
mating are not constrained by those of women. Suggestions for future
research in human mating psychology and behaviors were provided.

~

\emph{Keywords:} agent-based modeling, short-term mating, casual sex,
sexual strategy theory, female choice hypothesis

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Gender differences in sexual behaviors, especially short-term mating
behaviors, have been a well-documented phenomenon in human sexuality
research. For example, research has found that heterosexual men, as
compared to heterosexual women, have a higher average number of past
short-term sexual partners (Oliver \& Hyde, 1993; Petersen \& Hyde,
2010; Rissel et al., 2014), have short-term mating more frequently
(Petersen \& Hyde, 2010), and engage in extramarital sex more often
(Petersen \& Hyde, 2010). These gender differences have also been
observed between gay men and lesbian women (Bailey et al., 1994; Bryant
\& Demian, 1994; Peplau et al., 1997, 2004).

In the existing literature, gender differences in sexual behaviors are
not usually distinguished from those in attitudes towards or preferences
for short-term mating. Some researchers study gender differences on the
two levels simultaneously without a conceptual distinction (e.g.,
Petersen \& Hyde, 2010), whereas some implicitly equate the two,
assuming that behavioral differences are a direct expression of the
psychological ones (e.g., Schmitt et al., 2001).

Intuitive as this line of reasoning is, there are good reasons to doubt
its soundness. This is because any heterosexual sexual encounter
involves both a man and a woman (the frequency of encounters involving
more than two persons is negligible, e.g., Herbenick et al., 2017). A
new short-term mating experience for a man is, therefore, also a new
experience for a woman, and the same logic applies to counting a new
short-term mate. As a result, the psychological differences in
short-term mating preferences may not result in behavioral ones in the
heterosexual case (Archer, 2019).

The present study took gender differences in mating preferences as a
presumption, which was conceptualized by sexual strategy theory (Buss \&
Schmitt, 1993) and repeatedly found support by empirical studies (e.g.,
Schmitt, 2003; Walter et al., 2020). Using an agent-based model, we
investigated whether men and women's differential preferences for
short-term mating resulted in any gender differences in short-term
mating behaviors, specifically, in the number of short-term mating
experiences and short-term mates, and if they did, under what
circumstances.

\hypertarget{gender-differences-in-short-term-mating-preferences}{%
\subsection{Gender differences in short-term mating
preferences}\label{gender-differences-in-short-term-mating-preferences}}

Gender differences in mating preferences have been studied in the light
of sexual strategy theory (Buss \& Schmitt, 1993). It posits that since
the minimum investment that men devote to their offspring (contribution
of sperm through one sexual act) is lower than that of women (gestation,
labor, and lactation), men tend to have short-term mating as a larger
part of their mating strategy than women do. This is because women's
high minimum investment leads to higher opportunity cost (vs.~men) if
they have sex with a suboptimal partner and give birth to offspring with
a low survival chance. As a result, men may have a greater interest in
short-term mating, desire a larger number of short-term mates in a given
period, and be less selective with respect to accepting a potential mate
(Buss \& Schmitt, 1993).

The hypotheses derived from sexual strategy theory have received
extensive support from empirical studies. For example, men reported to
be currently seeking short-term mates to a larger extent than women did
(Buss \& Schmitt, 1993; Schmitt et al., 2001; Schmitt, 2003), and a
larger proportion of them were in any way seeking short-term mates
(vs.~not seeking) (Schmitt, 2003). These have been found among U.S.
college students (Buss \& Schmitt, 1993; Schmitt et al., 2001) as well
as in ten world regions in a cross-cultural study (Schmitt, 2003).
Similarly, men across the world also reported to desire more short-term
sexual partners compared to women for a number of given time periods
(e.g., a month, a year) (Buss \& Schmitt, 1993; McBurney et al., 2005;
Schmitt, 2003). The gender differences were significant regardless of
whether they were estimated by mean (Buss \& Schmitt, 1993; Schmitt et
al., 2001; Schmitt, 2003), median (Schmitt et al., 2001; Schmitt, 2003),
or percentage (Schmitt, 2003) statistics.

As for mating standards, men are less selective in accepting someone as
a potential short-term mate. For example, studies using U.S. college
samples found that the minimum percentile ranks that men found
acceptable for a potential short-term sexual partner were lower than
those accepted by women, both in terms of overall desirability and
individual traits (e.g., social status, attractiveness) (Kenrick et al.,
1990; Regan, 1998). A study also found that when presented with
identical descriptions of potential mates, men on average rated them as
more desirable than women did (Wiederman \& Dubois, 1998).

Some evidence shows that gender differences in mating preferences also
existed between gay men and lesbian women. A survey study using a
community sample from the U.S. found that gay men were more interested
in short-term mating than lesbian women were, and that this gender
difference was comparable to that existing among heterosexual
individuals (Bailey et al., 1994). An indirect piece of evidence on
differential standards for short-term mates comes from a recent study
finding that significantly more gay men than lesbian women reported to
have accepted a casual sexual offer from a same-gender person (Matsick
et al., 2021). Past studies suggest that the gender difference in the
acceptance rate of casual sexual offers may originate from men and
women's differential standards for short-term mates (Conley et al.,
2011; Hald \& Høgh-Olesen, 2010). Thus, as a postulation, gay men may
also have lower standards than lesbian women for short-term mates.

\hypertarget{constraints-on-mens-preferences-for-short-term-mating}{%
\subsection{Constraints on men's preferences for short-term
mating}\label{constraints-on-mens-preferences-for-short-term-mating}}

Since heterosexual sex involves both a man and a woman in most cases,
men's interest in short-term mating can be constrained by women's
preference for long-term relationships (Archer, 2019; Symons, 1979).
This is because men's short-term mating preferences can only translate
into behaviors when there are women willing to have sex with them. When
a new short-term mating encounter occurs, it counts towards men's total
number of short-term mating encounters as well as towards women's.
Similarly, a new short-term mate counts towards both men and women's
total number of short-term mates. Therefore, the total number of
short-term mating experiences and short-term mates must be equal between
heterosexual men and women on a population level. However, it is
important to note that although heterosexual men's total number of
short-term mating is equal to that of women, this does not necessarily
mean that the total number of men and women who have ever had short-term
mating must be equal. It is possible that there is a smaller proportion
of men (vs.~the proportion of women) who contribute to the total number
of short-term mating. Following this line of reasoning, we would expect
to observe no gender difference in short-term mating behaviors among
heterosexual individuals even when there were gender differences in
mating preferences.

As a comparison, in the cases of gay men and lesbian women, men's
preferences are not constrained by women's, but only by those of other
men, who, arguably, have more similar preferences. This would allow for
a more direct behavioral expression of men's mating preferences. The
notion that gay men have less restricted preferences was borne out in
the proportion of individuals who have engaged in extradyadic sex
(Peplau et al., 2004). A study showed that the proportions of
heterosexual men and women who have engaged in extradyadic sex were 26\%
and 21\%, respectively, while the statistics among gay men and lesbian
women were 82\% and 28\%, respectively (Peplau et al., 2004). Therefore,
we would theoretically expect to observe gender differences in
short-term mating behaviors between gay men and lesbian women. Moreover,
we would also expect to find that gay men, as compared to heterosexual
men, engaged in more short-term mating behaviors due to the lessened
constraint on their mating preferences.

\hypertarget{a-simple-model-of-short-term-mating-behaviors}{%
\subsection{A simple model of short-term mating
behaviors}\label{a-simple-model-of-short-term-mating-behaviors}}

By using a spatial agent-based model, the present study investigated
whether men and women's different preferences for short-term mating
resulted in any gender differences in short-term mating behaviors, and
if they did, under what circumstances. Preferences for short-term mating
were operationally defined as the interest in short-term mating and
standards for short-term mates (Buss \& Schmitt, 1993). Short-term
mating behaviors were operationally defined as the number of short-term
mating experiences and the number of past short-term mates (Petersen \&
Hyde, 2010).

We modeled this process among both heterosexual individuals and gay men
and lesbian women to examine whether constraints set by the opposite
sex's preferences would affect short-term mating behaviors. Individuals'
sexual orientation was conceptualized in terms of behaviors only in our
model. Heterosexual men had short-term mating or formed a long-term
relationship only with women, while gay men only with other men, and
vice versa for women.

We raised the following hypotheses in the present study: when there were
gender differences in preferences for short-term mating, 1) there would
be gender differences in short-term mating behaviors among gay men and
lesbian women, with gay men being engaged in more such behaviors, and 2)
gay men would engage in more short-term mating behaviors as compared to
heterosexual men.

\hypertarget{material-and-methods}{%
\section{Material and methods}\label{material-and-methods}}

\hypertarget{model-design}{%
\subsection{Model Design}\label{model-design}}

We developed an agent-based model to represent a simplified environment
where men and women can move around, search for mates, and form
long-term and/or short-term relationships. Space and time are modeled as
discrete variables. Space was represented as discrete locations on a
two-dimensional 33*33 lattice. Agents' movement in the space was not
meant to simulate physical movement but a state of encountering
potential mates. Staying at one location represented being committed to
a long-term relationship.

The model measured two outcomes: (1) the number of short-term mating
experiences of men and women; (2) the number of short-term mates of men
and women. Additionally, we also measured the number of men and women in
the mating pool (i.e., those with a non-zero number of short-term mating
experiences). The average numbers of short-term mating experiences and
short-term mates were calculated by taking the average among the whole
population of men/women (\emph{N\textsubscript{m}} =
\emph{N\textsubscript{w}} = 150) and among those who had engaged in
short-term mating.

See the Supplemental Materials for the overview, design concepts, and
details (ODD) protocol of the model, which includes detailed scheduling
and parameterization. The model can be downloaded from the Open Science
Framework {[}link masked for peer review{]}. The model was programmed in
Netlogo 6.2.1 (Wilensky, 1999).

\hypertarget{experiment-design}{%
\subsection{Experiment Design}\label{experiment-design}}

Using the agent-based model, the present study conducted two 2 (gender
difference vs.~no difference in the interest in short-term mating
likelihood) x 2 (gender difference vs.~no difference in mating
standards) experiments. Experiment 1 was run among heterosexual agents
who only engaged in short-term or long-term relationships with agents of
the opposite gender. Experiment 2 was run among gay men and lesbian
women who only engaged in short-term or long-term relationships with
agents of the same gender.

The interest in short-term mating was modeled as an agent's likelihood
of deciding to have short-term mating at each time step. The standards
for short-term mates were modeled as the minimum desirability of a
potential mate with whom an agent was willing to have sex. The values of
the two parameters were set based on empirical findings in human mating
psychology. When there was a gender difference in short-term mating
likelihood, men had a 40\% likelihood of deciding to engage in
short-term mating, while women had a likelihood of 25\% (Buss \&
Schmitt, 1993; Schmitt et al., 2001; Schmitt, 2003). When there was no
gender difference in short-term mating likelihood, both women and men
had a likelihood of 25\%. When there was a gender difference in mating
standards, men had a mating standard of 3 (the highest possible value =
10), while women had a standard of 5 (Kenrick et al., 1990; Regan,
1998). When there was no gender difference in mating standards, both men
and women had a standard of 5.

\hypertarget{model-schedule}{%
\subsection{Model Schedule}\label{model-schedule}}

\hypertarget{initial-setup}{%
\subsubsection{Initial setup}\label{initial-setup}}

In total, 150 women and 150 men were created on the lattice at random
locations. All agents were initialized with (1) a three-unit maximum
distance by which agents can move away from their birthplace (movement
range); (2) a 10\% likelihood of two agents forming a long-term
relationship upon meeting; (3) single status and no long-term partner;
(4) a mate value, sampled from a Gaussian distribution (\emph{M} = 5.0,
\emph{SD} = 1.5); (5) no short-term mate or short-term mating history.

The likelihood of engaging in short-term mating and the standard for
short-term mates were initialized either identically or differently for
men and women, depending on the experimental conditions.

\hypertarget{procedures}{%
\subsubsection{Procedures}\label{procedures}}

\hypertarget{heterosexual-procedures.}{%
\paragraph{Heterosexual procedures.}\label{heterosexual-procedures.}}

At each time step, agents first checked whether they were in a long-term
relationship. If they were not in a long-term relationship, they set
their heading randomly (if they were within the movement range from the
birthplace) or faced the birthplace (if they were out of the movement
range from the birthplace) and moved by a random distance. The random
distance was less than half of the movement range. If they were in a
long-term relationship, they did not move. Then, the agents decided on
whether to engage in short-term mating by a chance of the chosen
likelihood (either 25\% or 40\%).

Women checked to see if any men were at the same location. If there
were, they randomly chose one of them as their potential short-term
mate. If both men and women met each other's mating standards and both
of them decided to engage in short-term mating, they had sex. This would
result in increasing both parties' number of short-term mating
experiences by one. They also recorded each other on their lists of past
short-term mates if they were not on the lists yet.

Then, women randomly selected one man at the same location as their
potential long-term partner. There was a 10\% of chance that a pair
would form a long-term relationship. After forming a long-term
relationship, both men and women changed to coupled status and register
each other as the long-term partner.

\hypertarget{gay-men-and-lesbian-women-procedures.}{%
\paragraph{Gay men and lesbian women
procedures.}\label{gay-men-and-lesbian-women-procedures.}}

At each time step, men and women moved and decided on short-term mating
as the agents did in the heterosexual procedures. Half of the agents (75
men and 75 women) were also given the initiator status. The initiators
checked to see whether there were other agents at the same location. The
rest of the procedures were identical to the heterosexual procedures,
except that the agents only chose those with the same gender as
potential short-term mates or long-term partners.

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

The model was run for 1,000 time steps in each simulation. We ran 10,000
simulations for each experiment, including 2,500 simulations for each
condition. We controlled for initializing random seeds in the model
runs. All simulations were run using Netlogo 6.2.1 (Wilensky, 1999).

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

Statistical analyses were conducted using R version 4.1.3 (R Core Team,
2022) and the figures were generated by the ggplot2 package (Wickham,
2016). Two-tailed independent samples \emph{t}-tests were used for all
statistical comparisons. The data were assumed to be normally
distributed within each condition but was not formally tested.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{gender-differences-in-short-term-mating-behaviors}{%
\subsection{Gender differences in short-term mating
behaviors}\label{gender-differences-in-short-term-mating-behaviors}}

The results from Experiment 1 showed that when there were gender
differences in preferences for short-term mating, there was no gender
difference in the average number of short-term mating experiences among
heterosexual individuals (\emph{M\textsubscript{m}} = 1.43,
\emph{SD\textsubscript{m}} = 0.24; \emph{M\textsubscript{w}} = 1.43,
\emph{SD\textsubscript{w}} = 0.24), Cohen's \emph{d} = 0. Nor was there
a difference in the average number of short-term mates
(\emph{M\textsubscript{m}} = 0.80, \emph{SD\textsubscript{m}} = 0.11;
\emph{M\textsubscript{w}} = 0.80, \emph{SD\textsubscript{w}} = 0.11),
Cohen's \emph{d} = 0 (Figure 1).

\begin{figure}[h]
  \centering
  \includegraphics[width=70mm]{figures/fig1_men_vs_women.pdf}
  \caption{\textbf{Short-term mating behaviors of men and women after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables separately for heterosexual individuals and gay men and lesbian women. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of men and women in the model (\textit{N\textsubscript{m}} = \textit{N\textsubscript{w}} = 150).}
\end{figure}

The results from Experiment 2 showed that when there were gender
differences in preferences for short-term mating, there were gender
differences in short-term mating behaviors between gay men and lesbian
women (Figure 1). The average number of short-term mating experiences
was higher among gay men (\emph{M} = 2.63, \emph{SD} = 0.37) than among
lesbian women (\emph{M} = 0.32, \emph{SD} = 0.10), \emph{t}(4,998) =
-301.78, \emph{p} \textless{} .001, Cohen's \emph{d} = 8.54. Similarly,
the average number of short-term mates was higher among gay men
(\emph{M} = 1.72, \emph{SD} = 0.19) than among lesbian women (\emph{M} =
0.25, \emph{SD} = 0.07), \emph{t}(4,998) = -357.69, \emph{p} \textless{}
.001, Cohen's \emph{d} = 10.12.

In Experiment 1, however, when the means of the outcome variables were
calculated among men and women in the mating pool (the number of
individuals in the mating pool: \emph{M\textsubscript{m}} = 56.85,
\emph{SD\textsubscript{m}} = 5.99; \emph{M\textsubscript{w}} = 77.04,
\emph{SD\textsubscript{w}} = 7.53), there were gender differences in
short-term mating behaviors when men and women had differential
preferences for short-term mating. The average number of short-term
mating experiences was higher among heterosexual men (\emph{M} = 3.78,
\emph{SD} = 0.50) than among heterosexual women (\emph{M} = 2.79,
\emph{SD} = 0.35), \emph{t}(4,998) = -81.53, \emph{p} \textless{} .001,
Cohen's \emph{d} = 2.31. Likewise, the average number of short-term
mates was higher among heterosexual men (\emph{M} = 2.12, \emph{SD} =
0.17) than among heterosexual women (\emph{M} = 1.56, \emph{SD} = 0.11),
\emph{t}(4,998) = -135.85, \emph{p} \textless{} .001, Cohen's \emph{d} =
3.84.

\hypertarget{comparing-heterosexual-and-gay-men}{%
\subsection{Comparing heterosexual and gay
men}\label{comparing-heterosexual-and-gay-men}}

The results comparing across the two experiments revealed that gay men
engaged in short-term mating behaviors more than heterosexual men did
(Figure 2). The average number of short-term mating experiences was
higher among gay men (\emph{M} = 2.63, \emph{SD} = 0.37) than among
heterosexual men (\emph{M} = 1.43, \emph{SD} = 0.24), \emph{t}(4,998) =
-134.87, \emph{p} \textless{} .001, Cohen's \emph{d} = 3.81. Similarly,
the average number of short-term mates was higher among gay men
(\emph{M} = 1.72, \emph{SD} = 0.19) than among heterosexual men
(\emph{M} = 0.80, \emph{SD} = 0.11), \emph{t}(4,998) = -206.91, \emph{p}
\textless{} .001, Cohen's \emph{d} = 5.85.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/fig2_hetero_vs_gay_men.pdf}
  \caption{\textbf{Short-term mating behaviors of heterosexual and gay men after 1,000 time steps in the model when gender differences existed in mating preferences.} Violin plots summarizing the outcome variables. Plot (A) shows the average number of short-term mating experiences, and plot (B) shows the average number of short-term mates. Central points show mean values and whiskers represent standard errors, but the standard errors are small and overlap to form a single bar. Statistics were calculated using the full population of heterosexual and gay men in the model (\textit{N\textsubscript{hm}} = \textit{N\textsubscript{gm}} = 150).}
\end{figure}

\hypertarget{comparing-across-conditions}{%
\subsection{Comparing across
conditions}\label{comparing-across-conditions}}

In an exploratory manner, we also compared across conditions to see
which of the two dimensions of mating preferences contributed to gender
differences in short-term mating behaviors.

Among heterosexual individuals, gender differences in short-term mating
behaviors, as calculated among individuals in the mating pool, emerged
when men and women had different standards for short-term mates.
Heterosexual men (vs.~heterosexual women) in the mating pool had a
higher average number of short-term mating experiences and short-term
mates as long as they had lower mating standards, even when no gender
difference existed in short-term mating likelihood (experiences:
\emph{M\textsubscript{m}} = 2.82, \emph{SD\textsubscript{m}} = 0.36,
\emph{M\textsubscript{w}} = 2.21, \emph{SD\textsubscript{w}} = 0.28,
Cohen's \emph{d} = 1.89; partners: \emph{M\textsubscript{m}} = 1.76,
\emph{SD\textsubscript{m}} = 0.15; \emph{M\textsubscript{w}} = 1.38,
\emph{SD\textsubscript{w}} = 0.09, Cohen's \emph{d} = 3.13). In
comparison, virtually no gender differences in short-term mating
behaviors existed when men and women had the same mating standards
(Table 1).

\begin{table}

\caption{\label{tab:table_1}\textbf{Short-term mating behaviors of heterosexual men and women in the mating pool after 1,000 time steps in the model}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2in}llllllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{\makecell[c]{Average short-term mating\\experiences (in-pool)}} & \multicolumn{5}{c}{\makecell[c]{Average short-term\\mates (in-pool)}} \\
\cmidrule(l{3pt}r{3pt}){2-6} \cmidrule(l{3pt}r{3pt}){7-11}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
\multicolumn{1}{c}{\em{ }} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}}\\
\midrule
Same likelihood*Same standard & 2.21 & 0.38 & 2.20 & 0.37 & 0.03 & 1.39 & 0.12 & 1.39 & 0.12 & 0.04\\
Same likelihood*Different standard & 2.82 & 0.36 & 2.21 & 0.28 & 1.89 & 1.76 & 0.15 & 1.38 & 0.09 & 3.13\\
Different likelihood*Same standard & 2.80 & 0.48 & 2.78 & 0.47 & 0.03 & 1.57 & 0.14 & 1.56 & 0.14 & 0.05\\
Different likelihood*Different standard & 3.78 & 0.50 & 2.79 & 0.35 & 2.31 & 2.12 & 0.17 & 1.56 & 0.11 & 3.84\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\small
\item \textit{Note.} The statistics were calculated with subsamples of men and women who had had enagged in short-term mating behaviors. See Supplementary Materials for the mean and standard deviation of sample sizes.
\end{tablenotes}
\end{threeparttable}
\end{table}

Among gay men and lesbian women, gender differences in short-term mating
behaviors emerged either when they had different short-term mating
likelihood or when their mating standards were different. Gay men
(vs.~lesbian women) had a higher average number of short-term mating
experiences and short-term mates when they had a higher short-term
mating likelihood (experiences: \emph{M\textsubscript{m}} = 0.80,
\emph{SD\textsubscript{m}} = 0.23, \emph{M\textsubscript{w}} = 0.32,
\emph{SD\textsubscript{w}} = 0.10, Cohen's \emph{d} = 2.67; partners:
\emph{M\textsubscript{m}} = 0.52, \emph{SD\textsubscript{m}} = 0.13,
\emph{M\textsubscript{w}} = 0.25, \emph{SD\textsubscript{w}} = 0.07,
Cohen's \emph{d} = 2.60), or when they had lower mating standards
(experiences: \emph{M\textsubscript{m}} = 1.05,
\emph{SD\textsubscript{m}} = 0.18, \emph{M\textsubscript{w}} = 0.32,
\emph{SD\textsubscript{w}} = 0.10, Cohen's \emph{d} = 5.10; partners:
\emph{M\textsubscript{m}} = 0.82, \emph{SD\textsubscript{m}} = 0.12,
\emph{M\textsubscript{w}} = 0.25, \emph{SD\textsubscript{w}} = 0.07,
Cohen's \emph{d} = 5.78). In comparison, virtually no gender differences
in short-term mating behaviors existed when gay men and lesbian women
had the same short-term mating likelihood and mating standards (Table
2).

\begin{table}

\caption{\label{tab:table_2}\textbf{Short-term mating behaviors of gay men and lesbian women after 1,000 time steps in the model}}
\centering
\begin{threeparttable}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2in}llllllllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{\makecell[c]{Average short-term mating\\experiences}} & \multicolumn{5}{c}{Average short-term mates} \\
\cmidrule(l{3pt}r{3pt}){2-6} \cmidrule(l{3pt}r{3pt}){7-11}
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Men} & \multicolumn{2}{c}{Women} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){7-8} \cmidrule(l{3pt}r{3pt}){9-10}
\multicolumn{1}{c}{\em{ }} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{M}} & \multicolumn{1}{c}{\em{SD}} & \multicolumn{1}{c}{\em{d}}\\
\midrule
Same likelihood*Same standard & 0.32 & 0.10 & 0.32 & 0.10 & 0.04 & 0.25 & 0.07 & 0.25 & 0.07 & 0.03\\
Same likelihood*Different standard & 1.05 & 0.18 & 0.32 & 0.10 & 5.10 & 0.82 & 0.12 & 0.25 & 0.07 & 5.78\\
Different likelihood*Same standard & 0.80 & 0.23 & 0.32 & 0.10 & 2.67 & 0.52 & 0.13 & 0.25 & 0.07 & 2.60\\
Different likelihood*Different standard & 2.63 & 0.37 & 0.32 & 0.10 & 8.54 & 1.72 & 0.19 & 0.25 & 0.07 & 10.12\\
\bottomrule
\end{tabular}}
\begin{tablenotes}
\small
\item \textit{Note.} \textit{N\textsubscript{m}} = \textit{N\textsubscript{w}} = 150 in all conditions.
\end{tablenotes}
\end{threeparttable}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The present study aimed at using agent-based modeling to investigate
whether men and women's different preferences for short-term mating
resulted in any gender differences in short-term mating behaviors,
specifically, in the number of short-term mating experiences and
short-term mates, and if they did, under what circumstances. We raised
two hypotheses: 1) gay men would engage in more short-term mating
behaviors as compared to lesbian women, and 2) gay men would engage in
more short-term mating behaviors as compared to heterosexual men. The
results from 1,000 time steps in a model simulating men and women's
mating behaviors supported our hypotheses. First of all, as compared to
lesbian women, gay men had higher average numbers of short-term mating
experiences and short-term mates. Secondly, gay men also had higher
average numbers of short-term mating experiences and short-term mates as
compared to heterosexual men. In contrast, we found no gender
differences in short-term mating behaviors between heterosexual men and
women, although heterosexual men in the mating pool did have higher
average numbers of short-term mating experiences and short-term mates.

As we expected, heterosexual men and women did not differ in short-term
mating behaviors despite their differential preferences for short-term
mating. This was because heterosexual men and women had an equal total
number of short-term mating experiences , so as the total number of
short-term mates. Since the sex ratio was 1:1 in our model, the average
number of short-term mating experiences and short-term mates must be
equal between men and women as well. However, we did find that among
individuals in the mating pool, men engaged in more short-term mating
behaviors as compared to women. This was because there were more women
than men in the mating pool, resulting in lower averages among
heterosexual women despite the equal numbers of experiences and mates in
total.

Moreover, gender differences in short-term mating behaviors emerged when
heterosexual men and women had different mating standards, but not when
they had different short-term mating likelihood. When women had a higher
standard than men did, less men than women in the population were above
a potential partner's standard and thus had a chance to have sex with
them. This contributed to the unequal number of men and women in the
mating pool, which led to gender differences in short-term mating
behaviors among this population.

In the light of these results, the empirical observation of gender
differences in short-term mating behaviors among heterosexual
individuals (e.g., Petersen \& Hyde, 2010) appear to be perplexing
because this is illogical (e.g., Gurman, 1989). Possible explanations
for these empirical results have to highlight caveats in the observation
process. One possibility is that there was sampling bias in the
observations (e.g., Wiederman \& Dubois, 1998) since surveys regarding
short-term mating behaviors may tend to attract individuals who already
engage in such behaviors. Another possibility is that heterosexual men
and women tend to report short-term mating behaviors differently. This
difference may be a result of social desirability bias (Alexander \&
Fisher, 2003) or due to men and women's different estimation strategies
(e.g., men tend to round up) (Brown \& Sinclair, 1999).

Among gay men and lesbian women, large gender differences in short-term
mating behaviors existed when men and women had differential preferences
for short-term mating, which was in line with empirical observations
(e.g., Peplau et al., 1997, 2004). This was probably because the number
of short-term mating experiences and short-term mates no longer counted
towards men and women simultaneously. Any gender differences in mating
preferences would result in differences in behaviors. A closer look at
the results did support this postulation. Either a difference in the
likelihood of short-term mating or a difference in mating standards
alone could contribute to gender differences in mating behaviors. This
was perhaps because the former increased the probability of both parties
of a given gay couple deciding to have short-term mating, and the latter
increased the probability of both of them meeting each other's
standards, as compared to the case of a given lesbian couple.

Supporting the idea that gay men's preferences for short-term mating
were not constrained by those of women, we found that gay men engaged in
more short-term mating behaviors as compared to heterosexual men. This
was consistent with the empirical literature (e.g., Peplau et al.,
1997). Interestingly, gay men and heterosexual men had the same mating
likelihood and the same standards for short-term mates in our model. The
only difference was a change in the preferences of potential partners.
When men's partners had a stronger preferences for short-term mating
(i.e., having men vs.~having women as potential partners), men also
appeared to engage in more short-term mating behaviors.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Using agent-based modeling, the present study theoretically explored the
circumstances under which men and women's differential preferences for
short-term mating resulted in gender differences in short-term mating
behaviors. We found when men (vs.~women) had a stronger preferences for
short-term mating, heterosexual men and women engaged in short-term
mating behaviors to the same extent, while gay men engaged in more
short-term mating behaviors as compared to lesbian women. We also found
that gay men engaged in more short-term mating behaviors than
heterosexual men did. Theses results highlight the distinction between
preferences and behaviors in human mating. Individuals' mating behaviors
do not only depend on one's own preferences, but are also constrained by
partners' preferences. Future research in human mating should not only
focus on the psychological aspect but also pay attention to the
interaction between individuals' psychology and its context. These
results also cast doubt to the prevalent belief in the gender
differences in short-term mating behaviors, especially among
heterosexual individuals. Our findings suggest that there may be factors
in the observation process, such as sampling bias, that lead to the
observed differences. Future research in human sexuality should note
such possibilities and interpret any observed gender differences in
short-term mating behaviors cautiously.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

\hypertarget{data-availability}{%
\section{Data availability}\label{data-availability}}

All models, data, and analysis code can be downloaded at: {[}link masked
for peer review{]}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-alexander_truth_2003}{}}%
Alexander, M. G., \& Fisher, T. D. (2003). Truth and consequences: Using
the bogus pipeline to examine sex differences in self‐reported
sexuality. \emph{The Journal of Sex Research}, \emph{40}(1), 27--35.
\url{https://doi.org/10.1080/00224490309552164}

\leavevmode\vadjust pre{\hypertarget{ref-archer_reality_2019}{}}%
Archer, J. (2019). The reality and evolutionary significance of human
psychological sex differences. \emph{Biological Reviews}, \emph{94}(4),
1381--1415. \url{https://doi.org/10.1111/brv.12507}

\leavevmode\vadjust pre{\hypertarget{ref-bailey_effects_1994}{}}%
Bailey, J. M., Gaulin, S., Agyei, Y., \& Gladue, B. A. (1994). Effects
of gender and sexual orientation on evolutionarily relevant aspects of
human mating psychology. \emph{Journal of Personality and Social
Psychology}, \emph{66}(6), 1081--1093.
\url{https://doi.org/10.1037/0022-3514.66.6.1081}

\leavevmode\vadjust pre{\hypertarget{ref-brown_estimating_1999}{}}%
Brown, N. R., \& Sinclair, R. C. (1999). Estimating number of lifetime
sexual partners: Men and women do it differently. \emph{The Journal of
Sex Research}, \emph{36}(3), 292--297.
\url{https://doi.org/10.1080/00224499909551999}

\leavevmode\vadjust pre{\hypertarget{ref-bryant_relationship_1994}{}}%
Bryant, A. S., \& Demian. (1994). Relationship characteristics of
american gay and lesbian couples. \emph{Journal of Gay \& Lesbian Social
Services}, \emph{1}(2), 101--117.
\url{https://doi.org/10.1300/J041v01n02_06}

\leavevmode\vadjust pre{\hypertarget{ref-buss_sexual_1993}{}}%
Buss, D. M., \& Schmitt, D. P. (1993). Sexual strategies theory: An
evolutionary perspective on human mating. \emph{Psychological Review},
\emph{100}(2), 204--232.
\url{https://doi.org/doi:10.1037/0033-295X.100.2.204d}

\leavevmode\vadjust pre{\hypertarget{ref-conley_women_2011}{}}%
Conley, T. D., Moors, A. C., Matsick, J. L., Ziegler, A., \& Valentine,
B. A. (2011). Women, men, and the bedroom: Methodological and conceptual
insights that narrow, reframe, and eliminate gender differences in
sexuality. \emph{Current Directions in Psychological Science},
\emph{20}(5), 296--300. \url{https://doi.org/10.1177/0963721411418467}

\leavevmode\vadjust pre{\hypertarget{ref-s_j_gurman_six_1989}{}}%
Gurman, S. J. (1989). Six of one... \emph{Nature}, \emph{12}, 342.
\url{https://www.nature.com/articles/342012d0}

\leavevmode\vadjust pre{\hypertarget{ref-hald_receptivity_2010}{}}%
Hald, G. M., \& Høgh-Olesen, H. (2010). Receptivity to sexual
invitations from strangers of the opposite gender. \emph{Evolution and
Human Behavior}, \emph{31}(6), 453--458.
\url{https://doi.org/10.1016/j.evolhumbehav.2010.07.004}

\leavevmode\vadjust pre{\hypertarget{ref-herbenick_sexual_2017}{}}%
Herbenick, D., Bowling, J., Fu, T.-C. (Jane)., Dodge, B., Guerra-Reyes,
L., \& Sanders, S. (2017). Sexual diversity in the {United States}:
Results from a nationally representative probability sample of adult
women and men. \emph{{PLOS} {ONE}}, \emph{12}(7), e0181198.
\url{https://doi.org/10.1371/journal.pone.0181198}

\leavevmode\vadjust pre{\hypertarget{ref-kenrick_evolution_1990}{}}%
Kenrick, D. T., Sadalla, E. K., Groth, G., \& Trost, M. R. (1990).
Evolution, traits, and the stages of human courtship: Qualifying the
parental investment model. \emph{Journal of Personality}, \emph{58}(1),
97--116. \url{https://doi.org/10.1111/j.1467-6494.1990.tb00909.x}

\leavevmode\vadjust pre{\hypertarget{ref-matsick_gender_2021}{}}%
Matsick, J. L., Kruk, M., Conley, T. D., Moors, A. C., \& Ziegler, A.
(2021). Gender similarities and differences in casual sex acceptance
among lesbian women and gay men. \emph{Archives of Sexual Behavior},
\emph{50}(3), 1151--1166.
\url{https://doi.org/10.1007/s10508-020-01864-y}

\leavevmode\vadjust pre{\hypertarget{ref-mcburney_preferred_2005}{}}%
McBurney, D. H., Zapp, D. J., \& Streeter, S. A. (2005). Preferred
number of sexual partners: Tails of distributions and tales of mating
systems. \emph{Evolution and Human Behavior}, \emph{26}(3), 271--278.
\url{https://doi.org/10.1016/j.evolhumbehav.2004.09.005}

\leavevmode\vadjust pre{\hypertarget{ref-oliver_gender_1993}{}}%
Oliver, M. B., \& Hyde, J. S. (1993). Gender differences in sexuality: A
meta-analysis. \emph{Psychological Bulletin}, \emph{114}(1), 29--51.
https://doi.org/\url{http://dx.doi.org/10.1037/0033-2909.114.1.29}

\leavevmode\vadjust pre{\hypertarget{ref-peplau_national_1997}{}}%
Peplau, L. A., Cochran, S. D., \& Mays, V. (1997). A national survey of
the intimate relationships of african american lesbians and gay men: A
look at commitment, satisfaction, sexual behavior, and {HIV} disease. In
\emph{Ethnic and cultural diversity among lesbians and gay men} (pp.
11--38). Sage Publications.

\leavevmode\vadjust pre{\hypertarget{ref-peplau_sexuality_2004}{}}%
Peplau, L. A., Fingerhut, A., \& Beals, K. P. (2004). Sexuality in the
relationships of lesbians and gay men. In \emph{The handbook of
sexuality in close relationships} (pp. 349--369). Lawrence Erlbaum
Associates Publishers.

\leavevmode\vadjust pre{\hypertarget{ref-petersen_meta-analytic_2010}{}}%
Petersen, J. L., \& Hyde, J. S. (2010). A meta-analytic review of
research on gender differences in sexuality, 1993--2007.
\emph{Psychological Bulletin}, \emph{136}(1), 21--38.
\url{https://doi.org/10.1037/a0017504}

\leavevmode\vadjust pre{\hypertarget{ref-r}{}}%
R Core Team. (2022). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-regan_minimum_1998}{}}%
Regan, P. C. (1998). Minimum mate selection standards as a function of
perceived mate value, relationship context, and gender. \emph{Journal of
Psychology \& Human Sexuality}, \emph{10}(1), 53--73.
\url{https://doi.org/10.1300/J056v10n01_04}

\leavevmode\vadjust pre{\hypertarget{ref-rissel_heterosexual_2014}{}}%
Rissel, C., Badcock, P. B., Smith, A. M. A., Richters, J., Visser, R. O.
de, Grulich, A. E., Simpson, J. M., Rissel, C., Badcock, P. B., Smith,
A. M. A., Richters, J., Visser, R. O. de, Grulich, A. E., \& Simpson, J.
M. (2014). Heterosexual experience and recent heterosexual encounters
among {A}ustralian adults: The second {A}ustralian study of health and
relationships. \emph{Sexual Health}, \emph{11}(5), 416--426.
\url{https://doi.org/10.1071/SH14105}

\leavevmode\vadjust pre{\hypertarget{ref-schmitt_universal_2003}{}}%
Schmitt, D. P. (2003). Universal sex differences in the desire for
sexual variety: Tests from 52 nations, 6 continents, and 13 islands.
\emph{Journal of Personality and Social Psychology}, \emph{85}(1), 85.
\url{https://doi.org/10.1037/0022-3514.85.1.85}

\leavevmode\vadjust pre{\hypertarget{ref-schmitt_are_2001}{}}%
Schmitt, D. P., Shackelford, T. K., \& Buss, D. M. (2001). Are men
really more 'oriented' toward short-term mating than women? A critical
review of theory and research. \emph{Psychology, Evolution \& Gender},
\emph{3}(3), 211--239. \url{https://doi.org/10.1080/14616660110119331}

\leavevmode\vadjust pre{\hypertarget{ref-symons_evolution_1979}{}}%
Symons, D. (1979). \emph{The evolution of human sexuality}. Oxford
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-walter_sex_2020}{}}%
Walter, K. V., Conroy-Beam, D., Buss, D. M., Asao, K., Sorokowska, A.,
Sorokowski, P., Aavik, T., Akello, G., Alhabahba, M. M., Alm, C., Amjad,
N., Anjum, A., Atama, C. S., Atamtürk Duyar, D., Ayebare, R., Batres,
C., Bendixen, M., Bensafia, A., Bizumic, B., \ldots{} Zupančič, M.
(2020). Sex differences in mate preferences across 45 countries: A
large-scale replication. \emph{Psychological Science}, \emph{31}(4),
408--423. \url{https://doi.org/10.1177/0956797620904154}

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
Wickham, H. (2016). \emph{{ggplot2: Elegant Graphics for Data
Analysis}}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}

\leavevmode\vadjust pre{\hypertarget{ref-wiederman_evolution_1998}{}}%
Wiederman, M. W., \& Dubois, S. L. (1998). Evolution and sex differences
in preferences for short-term mates: Results from a policy capturing
study. \emph{Evolution and Human Behavior}, \emph{19}(3), 153--170.
\url{https://doi.org/10.1016/S1090-5138(98)00006-3}

\leavevmode\vadjust pre{\hypertarget{ref-wilensky_1999}{}}%
Wilensky, U. (1999). \emph{NetLogo}
{[}Http://ccl.northwestern.edu/netlogo/{]}. Center for Connected
Learning; Computer-Based Modeling.
\url{http://ccl.northwestern.edu/netlogo/}

\end{CSLReferences}

\end{document}