### 2022-06-10 submission/title_page.Rmd
added cover letter and title page, close #23

---
output:
  pdf_document
geometry: margin=1.0in
fontsize: 11pt
header-includes:
  - \newcommand{\bcenter}{\begin{center}}
  - \newcommand{\ecenter}{\end{center}}
  - \newcommand{\btitlepage}{\begin{titlepage}}
  - \newcommand{\etitlepage}{\end{titlepage}}
  - \usepackage{setspace}\doublespacing
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
---

\btitlepage

\bcenter

\vspace*{20mm}

# An Agent-Based Model of Gender and Sexual Orientation Differences in Short-Term Mating Behaviors as a Result of Mating Preferences {-}

\vspace{20mm}

Yurun Ying^a^^,^^$\dagger$^, Jan Antfolk^b^, Pekka Santtila^a^^,^^c^

^a^ Faculty of Arts and Sciences, NYU Shanghai, Shanghai, China

^b^ Faculty of Arts, Psychology and Theology, Åbo Akademi University, Turku, Finland

^c^ NYU-ECNU Institute for Social Development at NYU Shanghai, Shanghai, China

\ecenter

\vspace{30mm}

^$\dagger$^ To whom correspondence should be addressed: 1555 Century Avenue, Pudong New District, Shanghai, China, 200136. Email: [yurun.ying@nyu.edu](yurun.ying@nyu.edu).

\vspace{5mm}

**Declaration of Competing Interest:** We declare that there are no known competing interests that might have influenced the work reported in this paper.

\vspace{5mm}

Word count: 5103

\etitlepage